Computing in the Field of Rationals

### 2012-06-26 rational_ext.tex
_

  {{c v |--> v'}} iff {{<c, v, []> |-->* [c', v', [] ]}}
\end{proposition}
\begin{proof}
  To prove the above, we first show that a more general lemma holds,
  namely that: {{c v |--> v'}} iff {{<c, v, C> |-->* [c', v', C ]}}.

@@

%subcode{proof} include main
%@ c1 v1 |--> v1' ==> <c1, v1, C'> |-->* {[c1',v1',C']}
%@@ c1+c2 (left v1) |--> left v1' ==> ?

Choosing {{C' =Lft c2 C}}, we have the required derivation sequence: 
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1',v1', LeftP c2 C]} |--> {[c1'+c2,left v1',C]} }}. The proof follows similarly for the {{right v2}} case.

To show the right-to-left direction we proceed by induction of the
sequence of {{|-->*}} derivations. Again the case for primtive
isomorphisms follows trivially. Taking the case of {{c1+c2}} we are
required to show that given a sequence 
{{<c1+c2,v,C> |-->* {[c', v', C]} }} there is a derivation tree for
{{c1+c2 v |--> v'}}. In the case that {{v:v1+b2}} has the form
{{left~v1}}, this follows by observing that the given sequence must have the form
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1',v1', LeftP c2 C]} |--> {[c1'+c2,left v1',C]} }}. 
For the strictly smaller inner subsequence by induction we have
{{c1~v1 |--> v1'}} which lets us complete the derivation of
{{c1+c2~(left v1) |--> left v1'}}. The {{right v2}} follows similarly.

>>>>

  {{c v |--> v'}} iff {{<c, v, []> |-->* [c, v', [] ]}}
\end{proposition}
\begin{proof}
  To prove the above, we first show that a more general lemma holds,
  namely that: {{c v |--> v'}} iff {{<c, v, C> |-->* [c, v', C ]}}.

@@

%subcode{proof} include main
%@ c1 v1 |--> v1' ==> <c1, v1, C'> |-->* {[c1,v1',C']}
%@@ c1+c2 (left v1) |--> left v1' ==> ?

Choosing {{C' =Lft c2 C}}, we have the required derivation sequence: 
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1,v1', LeftP c2 C]} |--> {[c1+c2,left v1',C]} }}. The proof follows similarly for the {{right v2}} case.

To show the right-to-left direction we proceed by induction of the
sequence of {{|-->*}} derivations. Again the case for primtive
isomorphisms follows trivially. Taking the case of {{c1+c2}} we are
required to show that given a sequence 
{{<c1+c2,v,C> |-->* {[c1+c2, v', C]} }} there is a derivation tree for
{{c1+c2 v |--> v'}}. In the case that {{v:v1+b2}} has the form
{{left~v1}}, this follows by observing that the given sequence must have the form
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1,v1', LeftP c2 C]} |--> {[c1+c2,left v1',C]} }}. 
For the strictly smaller inner subsequence by induction we have
{{c1~v1 |--> v1'}} which lets us complete the derivation of
{{c1+c2~(left v1) |--> left v1'}}. The {{right v2}} follows similarly.
